---
title: "Exploring Linear Regression"
author: "JE Hug"
output: pdf_document
---

```{r setup, include=TRUE}
library(tidyverse)
library(furrr)
library(readxl)
library(magrittr)
```



```{r}

# initial clean of data
# read in prison data
prison <- read.csv(file = "./data/california_jail_survey_1995_2020_csv/california_jail_county_monthly_1995_2020.csv")
head(prison)

# read in revenue data
revenue <- read.csv(file = "./data/County_Revenues_Per_Capita.csv")


# read in GDP it is 2017 

calgdp <- read.csv("./data/gdpcali.csv",col.names = c("California", "GDP"))

calgdp$GDP <- as.numeric(gsub(",","",calgdp$GDP))


# filter revenue to 2017 and take out the counties that have no revenue
rev17 <- revenue %>% filter(Fiscal.Year == 2017) %>% filter(Entity.Name != "Alpine" & Entity.Name != "Sierra")

# filter prison to 2017 and take out countiees that have no prison
prison17 <- prison %>% filter(year == 2017) %>% filter(census_county_name !=  "San Francisco County")


avgpop <- prison17 %>% group_by(census_county_name) %>% summarise_at(vars(avg_daily_pop_sentenced_male,avg_daily_pop_sentenced_female,avg_felony_inmate_total),.funs = mean)


together <- cbind(calgdp,avgpop,rev17)

together %<>% mutate(gdppc = GDP/Estimated.Population) 

colnames(prison17)




```



```{r}
prisonvars <- prison17 %>% group_by(census_county_name) %>% 
  mutate(avg_daily_unsentenced = avg_daily_pop_unsentenced_female+avg_daily_pop_unsentenced_male) %>%   mutate(avg_daily_sentenced = avg_daily_pop_sentenced_female+avg_daily_pop_sentenced_male) %>% summarise_at(vars(avg_daily_unsentenced,avg_daily_sentenced,avg_felony_inmate_total,avg_fed_inmate_housed_contract),.funs = mean)

together <- cbind(calgdp,prisonvars,rev17)

together %<>% mutate(avg_daily_unsentenced_pc = avg_daily_unsentenced/Estimated.Population, 
                     avg_daily_sentenced_pc = avg_daily_sentenced/Estimated.Population,
                     avg_felony_inmate_total_pc = avg_felony_inmate_total/Estimated.Population,
                     avg_fed_inmate_housed_contract_pc = avg_fed_inmate_housed_contract/Estimated.Population,
                     gdppc = GDP/Estimated.Population) 

fit1 <- lm(Revenues.Per.Capita ~ avg_daily_unsentenced_pc + avg_daily_sentenced_pc, 
           data = together)


summary(fit1)

fit2 <- lm(Revenues.Per.Capita ~ avg_daily_unsentenced_pc + avg_daily_sentenced_pc+
             avg_felony_inmate_total_pc, 
           data = together)

summary(fit2)


fit3 <- lm(Revenues.Per.Capita ~  avg_daily_sentenced_pc+
             avg_felony_inmate_total_pc, 
           data = together)

summary(fit3)


# relationship between GDP and revenue 

fit4 <- lm(Revenues.Per.Capita~gdppc, data =together)

summary(fit4)

together %>% ggplot(aes(x=Revenues.Per.Capita,y=gdppc))+
  geom_point()+
  geom_smooth(method = "lm")

# the same fits with gdp

fit5 <- lm(log(gdppc) ~ avg_daily_unsentenced_pc + avg_daily_sentenced_pc, 
           data = together)

summary(fit5)

plot(fit5)


fit6 <- lm(log(gdppc) ~ avg_daily_unsentenced_pc + avg_daily_sentenced_pc+
             avg_felony_inmate_total_pc, 
           data = together)

summary(fit6)

fit7 <- lm(gdppc ~  avg_daily_sentenced_pc+
             avg_felony_inmate_total_pc, 
           data = together)

summary(fit7)


fit8 <- lm(gdppc~  avg_felony_inmate_total_pc, data = together)

summary(fit8)

plot(fit8)

# federal housing and revenue

fit9<- lm(Revenues.Per.Capita~avg_fed_inmate_housed_contract_pc,data = together)


summary(fit9)

together %>% ggplot(aes(x=avg_fed_inmate_housed_contract,y=Revenues.Per.Capita))+
  geom_point()+
  geom_smooth(method = "lm")


```



How many mental health case as a response where we have 
longitudinal data, as taking each measurement per month of a prison as a subject
each prison is a subject. We would need to adjust for population of prison somehow, 
the rate of mental health cases? 

```{r}

# make prisons factor
prison$jurisdiction <- as.factor(prison$jurisdiction)



# we can try an imputation approach

prisonMentH <- prison %>% mutate(mental_health_rate = mental_heath_case_open_end_month/highest_inmate_count) 


colnames(prisonMentH)



# lmer output

prisonMentH %<>% mutate(avg_daily_pop_unsentenced_male=scale(avg_daily_pop_unsentenced_male),avg_daily_pop_sentenced_male=scale(avg_daily_pop_sentenced_male)) 

lmef <- lme4::lmer(formula =  mental_health_rate ~ (1+ avg_daily_pop_unsentenced_male+avg_daily_pop_sentenced_male|jurisdiction),data = prisonMentH)

summary(lmef)

coefficients(lmef)

library(gee)

gef <- gee(formula = mental_health_rate ~ 1+ avg_daily_pop_unsentenced_male+avg_daily_pop_sentenced_male, id =jurisdiction, data = prisonMentH,corstr = "independence")

summary(gef)


library(geepack)

geepack::geeglm(formula = mental_health_rate ~ 1+ avg_daily_pop_unsentenced_male+avg_daily_pop_sentenced_male, id =jurisdiction, data = prisonMentH)
```

GEE vs LMER?

Mental health in prisons?


Regress County revenue on housing contracts

GO GET COUNTY LEVEL:

Poverty rate 
Crime rate
Median income 
Education level



Poverty data

```{r}
pov <- read_csv(file = "./data/povdata.txt",col_names = F)



x <- apply(pov,1,str_split,pattern = " ")


y <- unlist(x,recursive = FALSE)
y <- lapply(y, function(z){z[z!=""]})


povertydf <- data.frame(do.call(rbind,y))



cnames <- c("FIPS State Code", "FIPS county code","all ages in poverty","LB90CI all ages pov","UB90CI all ages pov",
              "% all ages", "LB90CI percent all ages pov","UB90CI percent all ages pov", 
              " 0-17 in poverty","LB90CI 0-17 pov","UB90CI 0-17 pov",
              "% 0-17", "LB90CI % 0-17 pov","UB90CI % 0-17 pov",
              "5-17 pov", "LB90CI children pov","UB90CI children pov","% 5-17", "LB90CI 5-17 % pov","UB90CI 5-17 % pov",
              "MHI", "90LB MHI", "90UB MHI")

colnames(povertydf) <- cnames

povertydf$`FIPS county code` <- as.numeric(povertydf$`FIPS county code`)
```

Current Goals:


1. Interpret all the simple OLS we did.
2. Use the extra covariates with Lasso and the SeletiveInference package
3. Use GEE to do across all the prison system, compare with multilevel model
4. What to do with other years?
Another GEE with multiple years, where each county is a group and we have an observation every year.

# FIRST FIX THE ONE DATASET BY GETTING IT A STUPID FIPS ROW

```{r RUN FIRST CHUNK TO RUN THIS}


countiesrevenues <-
  c("Alameda",
    "Amador",
    "Butte",
    "Calaveras",
    "Colusa",
    "Contra Costa",
    "Del Norte",
    "El Dorado",
    "Fresno",
    "Glenn",
    "Humboldt",
    "Imperial",
    "Inyo",
    "Kern",
    "Kings",
    "Lake",
    "Lassen",
    "Los Angeles",
    "Madera",
    "Marin",
    "Mariposa",
    "Mendocino",
    "Merced",
    "Modoc" ,
    "Mono"   ,
    "Monterey"     ,
    "Napa" ,
    "Nevada" ,
    "Orange"      ,
    "Placer"    ,
    "Plumas"    ,
    "Riverside" ,
    "Sacramento"     ,
    "San Benito",
    "San Bernardino" ,
    "San Diego"  ,
    "San Joaquin",
    "San Luis Obispo" ,
    "San Mateo"  ,
    "Santa Barbara",
    "Santa Clara" ,
    "Santa Cruz",
    "Shasta"    ,
    "Siskiyou"  ,
    "Solano"     ,
    "Sonoma",
    "Stanislaus"     ,
    "Sutter"  ,
    "Tehama"     ,
    "Trinity"      ,
    "Tulare"     ,
    "Tuolumne" ,
    "Ventura"    ,
    "Yolo"   ,
    "Yuba"
  )

countyFIPS <- sort(unique(prison17$fips_county_code))

countyFIPStable <- data.frame(Entity.Name = countiesrevenues,countyFIPS)



```







```{r}
# multiple years GEE part 4


# initial clean of data
# read in prison data
prison <- read.csv(file = "./data/california_jail_survey_1995_2020_csv/california_jail_county_monthly_1995_2020.csv")
head(prison)

# read in revenue data
revenue <- read.csv(file = "./data/County_Revenues_Per_Capita.csv")

# JOIN THE REVENUE WITH FIPS 

revenue <- inner_join(revenue,countyFIPStable, by = c("Entity.Name"))

revenue %<>% mutate(fips_county_code=countyFIPS.x)



# take out the counties that have no revenue
revall <- revenue  %>% filter(Entity.Name != "Alpine" & Entity.Name != "Sierra")%>% 
  filter(Fiscal.Year <= 2019 & Fiscal.Year >=2005) %>% mutate(year = Fiscal.Year)




# filter prison to other years and take out countiees that have no prison
prisonall <- prison %>% filter(census_county_name !=  "San Francisco County") %>% 
  filter(year <= 2019 & year >=2005) %>% arrange(census_county_name)


avgpop <- prisonall %>% group_by(fips_county_code,year) %>%
  summarise_at(vars(avg_daily_pop_sentenced_male,avg_daily_pop_sentenced_female,avg_felony_inmate_total,highest_inmate_count),.funs = mean)


together <- inner_join(avgpop,revall,by = c("fips_county_code","year"))


together %<>% mutate(avg_daily_pop_sentenced_male_pc = avg_daily_pop_sentenced_male/highest_inmate_count,
                   avg_daily_pop_sentenced_female_pc =  avg_daily_pop_sentenced_female/highest_inmate_count,
                    revPerPrisoner = Total.Revenues/highest_inmate_count)



```



```{r}
library(gee)
multi <- gee(revPerPrisoner ~ avg_daily_pop_sentenced_male_pc + avg_daily_pop_sentenced_female_pc,data = together, id =fips_county_code)

summary(multi)


library(lme4)
multilme <- lmer(revPerPrisoner ~ (1|fips_county_code) +avg_daily_pop_sentenced_male_pc + avg_daily_pop_sentenced_female_pc, data = together)

summary(multilme)


multilmersl <- lmer(revPerPrisoner ~ (1 +avg_daily_pop_sentenced_male_pc |fips_county_code), data = together)
summary(multilmersl)

```


